Pure Spin Current Injection in Hydrogenated Graphene Structures

### 2017-07-28 mu/article/article.tex
Starting review 5
New files with 0.0002 of broadening.
Including abstract and almost final modifications.

\title{Pure Spin Current Injection in Hydrogenated Graphene Structures}
\author{Reinaldo Zapata-Pe\~na\textsuperscript{1},
        Bernardo S. Mendoza\textsuperscript{1},
        Anatoli I. Shkrebtii\textsuperscript{2}}
\affiliation{\textsuperscript{1}Centro de Investigaciones en \'Optica, Le\'on,
Guanajuato 37150, M\'exico}
\affiliation{\textsuperscript{2}University of Ontario, Institute of Technology,
Oshawa, ON, L1H 7L7, Canada}

@@

\begin{abstract}
perro
%\blindtext
\end{abstract}

@@

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\linewidth]{figures/upstruc2}
    \\
    \includegraphics[width=\linewidth]{figures/upstruc1}
    \caption{Up structure}
    \label{fig:up-struc}
\end{figure}
\begin{figure}[ht!]
    \centering
    \includegraphics[width=\linewidth]{figures/altstruc2}
    \\
    \includegraphics[width=\linewidth]{figures/altstruc1}
    \caption{Alt structure.}
    \label{fig:alt-struc}
\end{figure}

Spintronics is an emerging research field of electronics in which the
manipulation and transport of spin of electrons in a solid state 
media plays the determining role adding a new degree of freedom to the
conventional charge manipulation.\cite{wolfSC04,fabianAPS07}
% 
At present there is an increasing interest in attain the same level of control
over the transport of spin at micro or nano scales as has been done for the
flow of charge in typical electronic devices.\cite{awschalomNP2007} Some
semiconductor spintronics devices have been proposed \cite{majumdarAPL06,
dattaAPL90,gotteNat16,pershinPRB08} and some of them require spin polarized
electrical current \cite{awschalomSSBM13} or pure spin current (PSC).
% 
One of the difficulties to achieve the development of spin current and PSC
semiconductor devices is the fact that the spin relaxation time in a
semiconducting media is short disabling the spin transport and then resulting
in a no observable spin current.\cite{murakamiSc03}
% 
In PSCs there is no net
motion of charge; spin-up electrons move in a given direction while spin-down
electrons travel in the opposite one. This phenomena can result from spin
injection,\cite{malPRB03} Hall Effects,\cite{sinovaPRB04} interference of two
optical beams,\cite{bhatPRL00, najmaiePRB03} or one photon absorption of
linearly polarized light\cite{bhatPRL05} and has been observed in gallium
arsenide (GaAs),\cite{zhaoPRL2006, stevensPRL03} aluminum-gallium arsenide
(AlGaAs),\cite{stevensPRL03} and Co$_2$FeSi.\cite{kimuraNGPAM12}

Graphene, an allotrope of carbon with hexagonal 2D lattice structure presents
properties like fractional quantum Hall effect at room temperature, excellent
thermal transport properties, excellent conductivity\cite{heerscheNat07} and
strength \cite{geimNM07, reinaNL08, novoselov2S07, balandinNL08} being then a
perfect platform to be used in two-dimensions electronic systems; however most
electronic applications are disabled by the absence of a semiconducting gap.
Recent studies demonstrate that the band gap of graphene can be opened by
applying an electric field,\cite{zhangN09} reducing the surface
area,\cite{hanPRL07} or applying uniaxial strain.\cite{niACSN08} Another
possibility to open the gap is by doping; this has been successfully achieved
using nitrogen,\cite{weiNL2009} boron-nitrogen,\cite{guoIJ11}
silicon,\cite{colettiPRB10} noble-metals,\cite{varykhalovPRB10} and
hydrogen.\cite{eliasS09, guisingerNL09, samarakoonACSN10}
% 
Depending on the percentage of hydrogenation and spatial configurations of
hydrogen-carbon bonds, hydrogenated graphene can result in different spatial
configurations.
% 
In this paper we present two 50\% hydrogenated graphene noncentrosymmetric
structures both presenting a discernible band gap: the \emph{Up} structure,
shown in Fig. \ref{fig:up-struc}, has hydrogen atoms bonded to the carbon layer
only in the upper side of the structure while the \emph{Alt} structure, shown
in Fig. \ref{fig:alt-struc}, has hydrogen alternating in the upper and bottom
sides of the carbon slab.\cite{zapataPSB2016}

Using those structures we address a theoretical study of the spin velocity
injection (SVI) by one-photon absorption of linearly polarized light.
% 
% We calculated the responses for the particular cases when the spin of
% electrons is directed along the the $z$ Cartesian coordinate, perpendicular
% to the $xy$ plane of the structure, or for the case when the velocity is
% directed along the $x$ and $y$ Cartesian direction on the $xy$ plane of the
% structures.
{
% \color{red}
% 
Because we have 2D structures we made the analysis for tow cases. The first is
fixing the spin of the electrons along the $z$ Cartesian direction with the
velocity directed on the surface of the structure in the $xy$ plane. The second
is fixing the spin velocity in the $x$ or $y$ direction and the spin directed
in $xyz$.}
% 
The SVI is an optical effect that quantifies the velocity at which a PSC moves
along the Cartesian direction $\mathrm{a}$ with the spin of electron polarized
along the Cartesian direction $\mathrm{b}$. One photon absorption of linearly
polarized light can promote a distribution of electrons in $\mathbf{k}$ space
regardless the symmetry of the material resulting in a null electrical current.
Then, the electrons excited to the conduction bands at opposite $\mathbf{k}$
points will result in opposite spin polarizations producing no net spin
injection.\cite{bhatPRL05} If the crystalline structure of the material is
noncentrosymmetric the spin polarization injected at a given $\mathbf{k}$ point
could not vanish\cite{alvaradoPRL85, schmiedeskampPRL88} 
% 
{
% \color{red}
% 
and then, since the velocities of electrons at opposite $\mathbf{k}$ points are
opposite, a pure spin current will be produced. Since the structures presented
here are noncentrosymmetric, they are good candidates in which this effect can
be induced.}

This paper is organized as follows. In Section \ref{sec:theory} we present the
theory and formulas that describe PSC and SVI. In Section \ref{sec:results} we
describe the details of calculations and the corresponding SVI spectra for the
\emph{Up} and \emph{Alt} structures. Finally, we present our conclusions in
Section \ref{sec:conclusions}.

@@

In this section, we report a summary of the theory that involves the PSC
phenomena from which rises the SVI treated in this paper.

{\color{red}
In PSCs there is no net motion of electrical charge and spin-up electrons move
in a given direction while spin-down electrons travel in the opposite one.
This effect can result from one photon absorption of linearly polarized light
by a semiconductor, with filled valence bands and empty conduction bands,
illuminated by light with photon energy larger than the energy gap.
% note 1
Using a single continuous linearly polarized laser beam, it is possible
to promote electrons in $\mathbf{k}$ space regardless the symmetry of the
system resulting in a net current equal to zero. 
% 
If the phenomena is produced in a noncentrosymmetric semiconducting media, the
electrons promoted to the conduction bands at opposite $\mathbf{k}$ points
can produce a net spin polarization different to zero\cite{alvaradoPRL85},
resulting in a PSC because the velocity of electrons
at opposite $\mathbf{k}$ points are in opposite directions.
}

@@

The operator that describes the electronic PSC is written as

@@

is the pseudotensor that describes the rate of change of the  PSC process in semiconductors.
To derive above we used  

@@

that follows from time-reversal invariance. 
The prime $'$ in the sum means that $c$ and $c'$ are quasi degenerate states and the
sum only covers these states.  Since $\mu^{\mathrm{abcd}}(\omega)$ is real we
have that $\mu^{\mathrm{abcd}}(\omega) =

@@

The function $\calv^{\rma\rmb}(\go)$ allow us to quantify two very
important aspects of PSC. On one hand, we can fix the spin direction
along $\mathrm{b}$,   
and calculate the resulting electron velocity.  
On the other hand, we can fix de velocity of the electron
along $\mathrm{b}$,  and study the resulting direction along which the
spin is polarized.
To this end, the added advantage of  2D structures, besides choosing
them noncentrosymmetric, is that we can use an incoming linearly
polarized beam of light
at normal incidence, and use the  direction of the polarized  electric
field to control $\calv^{\rma\rmb}(\go)$.
Indeed, writing 
$\bfE(\omega) = E_0(\omega)(\cos\ga\,\hat\bfx+\sin\ga\,\hat\bfy)$
where $\alpha$ is the polarization angle, we obtain from
 Eq. \eqref{eq:vab-w}
that
\begin{widetext}
\begin{align}
\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)
&= 
\frac{2}{\hbar\xi(\go)}
\left(\mu^{\mathrm{abxx}}(\omega)\cos^{2}\alpha + 
\mu^{\mathrm{abyy}}(\omega)\sin^{2}\alpha + 
\mu^{\mathrm{abxy}}(\omega)\sin 2\alpha\right)
\label{eq:vab-aw}
\end{align}
\end{widetext}
as for the structures chosen in this article,
$\xi^{\mathrm{xx}}(\go)=\xi^{\mathrm{yy}}(\go)\equiv\xi(\go)$, and $\xi^{\mathrm{xy}}(\go)=0$.
Now, we formalize our two options for $\calv^{\rma\rmb}$.
%

% Two interesting possibilities to analyze the SVI are fixing the spin along $z$,
% directed perpendicularly to the surface of the structure, or fixing the
% velocity along $x$ or $y$ on the $xy$ plane of the structures. Also we can
% analyze the SVI contribution coming from each layer of the structure. In
% following subsections we present these three cases.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%% Theory: Fixing spin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Analyzing the SVI, Eq. \eqref{eq:vab-aw}, we define the magnitude of
the electron's in plane
velocity with it's spin polarized along the $\mathrm{b}$ direction as
\begin{equation}
\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)
\equiv
\sqrt{
[\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)]^{2}\ +
[\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)]^{2}\ 
}, 
\label{eq:vs-mag}
\end{equation}
and define the angle at which the velocity is directed on the $xy$ plane as
\begin{equation}
\gamma_{\gs^\mathrm{b}} (\omega,\alpha)
=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)} \right)
.
\label{eq:gamma-ang}
\end{equation}
% where this angle is measured in the counter-clockwise direction from the
% positive $x$. 
We also define two special angles
\begin{equation}
\gamma_{\gs^\mathrm{b}}^\parallel(\omega,\alpha) = \alpha, 
\label{eq:gamma-par} 
\end{equation}
and
\begin{equation}
\gamma_{\gs\mathrm{b}}^\perp(\omega,\alpha) = \alpha \pm 90^{\circ},
\label{eq:gamma-perp}
\end{equation}
corresponding to the electron velocity being parallel or perpendicular
the incoming polarization, 
respectively. The subscript $\gs^\rmb$ denotes the spin along $\rmb$.

@@

\begin{table}[t]
\center
\begin{tabular}{ccccc}\\
\hline
\quad Layer \quad & \quad Atom \qquad & \multicolumn{3}{c}{Position [\AA]} \\
\cline{3-5}
\quad No.   \quad & \quad type \qquad & $x$ & $y$ & $z$  \\
\hline
1 & H & -0.61516 & -1.77416 &  0.73196 \\
1 & H &  0.61518 &  0.35514 &  0.73175 \\
2 & C & -0.61516 & -1.77264 & -0.49138 \\
2 & C & -0.61516 & -0.35600 & -0.72316 \\
2 & C &  0.61516 &  0.35763 & -0.49087 \\
\hline
\end{tabular}

\caption{Unit cell of \emph{Up} structure. Layer division, atom types and
positions for the \emph{Up} structure. The structure unit cell was divided in
two layers corresponding to hydrogen and carbon atoms.The corresponding layer
atom position is depicted in Fig. \ref{fig:up-struc} with the corresponding
number of layer.}
\label{tab:up-unitcell}
\end{table}
% 
% 
\begin{table}[t]
\center
\begin{tabular}{ccccc}\\
\hline
\quad Layer \quad & \quad Atom \qquad & \multicolumn{3}{c}{Position [\AA]} \\
\cline{3-5}
\quad No.   \quad & \quad type \qquad & $x$ & $y$ & $z$  \\
\hline
1 & H &  -0.61516 &  -1.42140 & \ 1.47237 \\
2 & C &  -0.61516 &  -1.73300 & \ 0.39631 \\
3 & C & \ 0.61516 & \ 1.73300 & \ 0.15807 \\
4 & C & \ 0.61516 & \ 0.42201 &  -0.15814 \\
5 & C &  -0.61516 &  -0.37396 &  -0.39632 \\
6 & H &  -0.61516 &  -0.68566 &  -1.47237 \\
\hline
\end{tabular}

\caption{Unit cell of \emph{Alt} structure. Layer division, atom types and
positions for the \emph{Alt} structure. The structure unit cell was divided in
six layers corresponding each one to atoms in different $z$ positions. The
corresponding layer atom position is depicted in Fig. \ref{fig:alt-struc} with
the corresponding number of layer.}
\label{tab:alt-unitcell}
\end{table}

We preset the results for 
and
$\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)$ 
and
$\mathcal{V}_{\mathrm{a}}(\omega,\alpha)$ 
for the
C$_{16}$H$_{8}$-up  and C$_{16}$H$_{8}$-alt 
structures being both
noncentrosymmetric semi-infinite 2D carbon systems with 50\% hydrogenation in
different arrangements. We recall that the \emph{Up} structure has hydrogen
atoms only on the upper side of the carbon sheet while the \emph{Alt} structure
has alternating hydrogen atoms on the upper and bottom sides. Also we take the
hexagonal carbon lattice to be on the $xy$ plane for both structures, and the
carbon-hydrogen bonds on the perpendicular $xz$ plane, as depicted in Figs.
\ref{fig:up-struc} and \ref{fig:alt-struc}. The coordinates for the
\emph{Up} and \emph{Alt} unit cells of the structures are presented in Tables
\ref{tab:up-unitcell} and \ref{tab:alt-unitcell}.

We calculated the self-consistent ground state and the Kohn-Sham
states using density functional theory in the local density approximation 
(DFT-LDA) with a planewave basis using the ABINIT code \cite{gonzeCPC09}.
% 
We used Hartwigsen-Goedecker-Hutter (HGH) relativistic separable dual-space
Gaussian pseudopotentials \cite{hartwigsenPRB98} including the spin-orbit
interaction needed to calculate $\mu^{\mathrm{abcd}}(\omega,\alpha)$ presented
in Eq. \eqref{eq:mu}.
% 
The convergence parameters for the calculations of our results corresponding to
the \emph{Up} and \emph{Alt} structures are cutoff energies of 40\,Ha and
65\,Ha, resulting in LDA energy band gaps of 0.084\,eV and 0.718\,eV,
respectively. The energy eigenvalues and matrix elements for the \emph{Up} and
\emph{Alt} structures were calculated using 12802 $\mathbf{k}$ points and 14452
$\mathbf{k}$ points in the IBZ to integrate $\mu^{\mathrm{abcd}}(\go)$ and
$\xi^{\rma\rmb}(\go)$ using the linearized analytic tetrahedron method
(LATM).\cite{nastosPRB07} We neglect the anomalous velocity term
$\hbar(\bfgs\times\nabla V)/4m^2c^2$, where $V$ is the crystal potential, in
$\hat\bfv$ of Eq.~\eqref{z.1}, as this term is known to give small contribution
to PSC.\cite{bhatPRL05} Therefore, $[\hat\bfv,\hat\bfS]=0$ and Eq.~\eqref{z.1}
reduces to $\hat K^{\rma\rmb}=\hat v^\rma \hat S^\rmb=\hat S^\rmb \hat v^\rma$.

% 
% It is known that the DFT with LDA approximation misestimates the energy band
% gap of semiconductors. To correct this one has to include the many-body
% interaction using i.e. the so-called GW approximation but this technique has a
% very high computational cost and is out of the scope in this paper.
% Nevertheless, DFT still remains a mainstream tool for calculating diverse
% properties derived from the electronic band structure.

@@

\begin{table}[b]
\begin{tabular}{cccccc}
\hline
\multirow{2}{*}{Structure \quad} & 
Kind of \quad & 
Pol. &
Energy & 
\multicolumn{2}{c}{$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$}\\
\cline{5-6}
& system & Ang. & [eV] & $\mathrm{ab}$ \quad & [Km/s]\\
\hline
\emph{Up}    & 2D   & 35    & 0.084  & $\mathrm{yz}$ &  684.2    \\
             &      &       & 1.954  & $\mathrm{yz}$ &  293.9    \\
             &      &       & 1.958  & $\mathrm{yz}$ & -273.0    \\
\emph{Alt}   & 2D   & 150   & 0.720  & $\mathrm{yz}$ & -722.2    \\
             &      &       & 0.911  & $\mathrm{yz}$ & -351.7    \\
 CdSe        & bulk & 90    & 0.912  & $\mathrm{zz}$ &  -32.3    \\
 GaAs        & bulk & 90    & 2.326  & $\mathrm{xx}$ &  -28.9    \\
\hline
\end{tabular}

\caption{Comparison of the reported maxima values of
$\mathcal{V}^{\mathrm{ab}}$ for the different structures and their
corresponding polarization angle $\alpha$ and energy values of the incoming
beam at which the maxima is obtained.}
\label{tab:vab-str-comp}
\end{table}

>>>>

\title{Pure Spin Current Injection in Hydrogenated Graphene Structures}
\author{Reinaldo Zapata-Pe\~na and
        Bernardo S. Mendoza}
\affiliation{Centro de Investigaciones en \'Optica, Le\'on,
Guanajuato 37150, M\'exico}
\author{Anatoli I. Shkrebtii}
\affiliation{University of Ontario, Institute of Technology,
Oshawa, ON, L1H 7L7, Canada}

@@

\begin{abstract}
We present a theoretical study of spin-velocity injection with linearly
polarized light of two 50\% hydrogenated noncentrosymmetric graphene structures:
Alt and Up. We analyzed the cases when the spin is polarized perpendicularly to
the surface of the structures, in the $z$ direction, and wen the spin moves
along the fixed $x$ and $y$ directions. The results are calculated in a full
electronic band structure scheme within DFT in the LDA approximations showing an
anisotropic behavior. We obtained maximum absolute values of
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha) = 668.0$\,Km/s for the Up
structure and $\mathcal{V}^{\mathrm{y}} (\omega,\alpha) = 905.6$\,Km/s for the
Alt structure being both structures excellent candidates for spintronics
applications.
\end{abstract}

@@

Spintronics is an emerging research field of electronics in which the
manipulation and transport of spin of electrons in a solid state 
media plays the determining role adding a new degree of freedom to the
conventional charge manipulation.\cite{wolfSC04,fabianAPS07}
% 
At present there is an increasing interest in attain the same level of control
over the transport of spin at micro or nano scales as has been done for the
flow of charge in typical electronic devices.\cite{awschalomNP2007} Some
semiconductor spintronics devices have been proposed \cite{majumdarAPL06,
dattaAPL90,gotteNat16,pershinPRB08} and some of them require spin polarized
electrical current \cite{awschalomSSBM13} or pure spin current (PSC).
% 
One of the difficulties to achieve the development of spin current and PSC
semiconductor devices is the fact that the spin relaxation time in a
semiconducting media is short disabling the spin transport and then resulting
in a no observable spin current.\cite{murakamiSc03}
% 
In PSCs there is no net
motion of charge; spin-up electrons move in a given direction while spin-down
electrons travel in the opposite one. 
This effect can result from one photon absorption of linearly polarized light
by a semiconductor, with filled valence bands and empty conduction bands,
illuminated by light with photon energy larger than the energy gap.
For instance,
this phenomena can result from spin
injection,\cite{malPRB03} Hall Effects,\cite{sinovaPRB04} interference of two
optical beams,\cite{bhatPRL00, najmaiePRB03} or one photon absorption of
linearly polarized light\cite{bhatPRL05} and has been observed in gallium
arsenide (GaAs),\cite{zhaoPRL2006, stevensPRL03} aluminum-gallium arsenide
(AlGaAs),\cite{stevensPRL03} and Co$_2$FeSi.\cite{kimuraNGPAM12}

Graphene, an allotrope of carbon with hexagonal 2D lattice structure presents
properties like fractional quantum Hall effect at room temperature, excellent
thermal transport properties, excellent conductivity\cite{heerscheNat07} and
strength \cite{geimNM07, reinaNL08, novoselov2S07, balandinNL08} being then a
perfect platform to be used in two-dimensions electronic systems; however most
electronic applications are disabled by the absence of a semiconducting gap.
Recent studies demonstrate that the band gap of graphene can be opened by
applying an electric field,\cite{zhangN09} reducing the surface
area,\cite{hanPRL07} or applying uniaxial strain.\cite{niACSN08} Another
possibility to open the gap is by doping; this has been successfully achieved
using nitrogen,\cite{weiNL2009} boron-nitrogen,\cite{guoIJ11}
silicon,\cite{colettiPRB10} noble-metals,\cite{varykhalovPRB10} and
hydrogen.\cite{eliasS09, guisingerNL09, samarakoonACSN10}
% 
Depending on the percentage of hydrogenation and spatial configurations of
hydrogen-carbon bonds, hydrogenated graphene can result in different spatial
configurations.
% 
In this paper we present two 50\% hydrogenated graphene noncentrosymmetric
structures both presenting a discernible band gap: the Up structure,
shown in Fig. \ref{fig:up-struc}, has hydrogen atoms bonded to the carbon layer
only in the upper side of the structure while the Alt structure, shown
in Fig. \ref{fig:alt-struc}, has hydrogen alternating in the upper and bottom
sides of the carbon slab.\cite{zapataPSB2016}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\linewidth]{figures/upstruc2}
    \\
    \includegraphics[width=\linewidth]{figures/upstruc1}
    \caption{Side (top panel) and top (bottom panel) views of the Up
      structure along with the 
      Cartesian $xyz$ directions. The dark (light) spheres are the C (H) atoms,
      labeled 1 (2).
}
    \label{fig:up-struc}
\end{figure}
\begin{figure}[ht!]
    \centering
    \includegraphics[width=\linewidth]{figures/altstruc2}
    \\
    \includegraphics[width=\linewidth]{figures/altstruc1}
    \caption{Side (top panel) and top (bottom panel) views of the Alt
      structure along with the 
      Cartesian $xyz$ directions. The dark (light) spheres are the C (H) atoms,
      labeled 1 (2).
}
    \label{fig:alt-struc}
\end{figure}

Using those structures we address a theoretical study of the 
spin velocity
injection (SVI) 
by one-photon absorption of linearly polarized light.
% 
% We calculated the responses for the particular cases when the spin of
% electrons is directed along the the $z$ Cartesian coordinate, perpendicular
% to the $xy$ plane of the structure, or for the case when the velocity is
% directed along the $x$ and $y$ Cartesian direction on the $xy$ plane of the
% structures.
{
% \color{red}
% 
Because we have 2D structures we made the analysis for tow cases. The first is
fixing the spin of the electrons along the $z$ Cartesian direction with the
velocity directed on the surface of the structure in the $xy$ plane. The second
is fixing the spin velocity in the $x$ or $y$ direction and the spin directed
in $xyz$.}
% 
The SVI is an optical effect that quantifies the velocity at which a PSC moves
along the Cartesian direction $\mathrm{a}$ with the spin of electron polarized
along the Cartesian direction $\mathrm{b}$. One photon absorption of linearly
polarized light can promote an even distribution of electrons in $\mathbf{k}$ space
regardless the symmetry of the material resulting in a 
null electrical current.\cite{bhatPRL05}
Then, the electrons excited to the conduction bands at opposite $\mathbf{k}$
points will result in opposite spin polarizations producing no net spin
injection.\cite{bhatPRL05} If the crystalline structure of the material is
noncentrosymmetric the spin polarization injected at a given $\mathbf{k}$ point
not necessarily 
 vanishes,\cite{alvaradoPRL85, schmiedeskampPRL88} 
and then, since the velocities of electrons at opposite $\mathbf{k}$ points are
opposite, a PSC will be produced. Since the structures presented
here are noncentrosymmetric, they are good candidates in which this effect can
be induced.

This paper is organized as follows. In Section \ref{sec:theory} we present the
theory and formulas that describe PSC and SVI. In Section \ref{sec:results} we
describe the details of calculations and the corresponding SVI spectra for the
Up and Alt structures. Finally, we present our conclusions in
Section \ref{sec:conclusions}.

@@

In this section, we report a summary of the theory  involved in the
calculation of the spin velocity
injection (SVI) resulting from the
pure spin current (PSC).

@@

The operator that describes the electronic SVI is written as

@@

is the pseudotensor that describes the rate of change of the  PSC process in
semiconductors. To derive above we used

@@

that follows from time-reversal invariance. The prime $'$ in the sum means that
$c$ and $c'$ are quasi degenerate states and the sum only covers these states.
Since $\mu^{\mathrm{abcd}}(\omega)$ is real we have that
$\mu^{\mathrm{abcd}}(\omega) =

@@

The function $\calv^{\rma\rmb}(\go)$ allow us to quantify two very
important aspects of PSC. On one hand, we can fix the spin direction
along $\mathrm{b}$,   
and calculate the resulting electron velocity.  
On the other hand, we can fix de velocity of the electron
along $\mathrm{b}$,  and study the resulting direction along which the
spin is polarized.
To this end, the added advantage of  2D structures, besides choosing
them noncentrosymmetric, is that we can use an incoming linearly
polarized beam of light
at normal incidence, and use the  direction of the polarized  electric
field to control $\calv^{\rma\rmb}(\go)$.
Indeed, writing 
$\bfE(\omega) = E_0(\omega)(\cos\ga\,\hat\bfx+\sin\ga\,\hat\bfy)$
where $\alpha$ is the polarization angle, we obtain from
 Eq. \eqref{eq:vab-w}
that
\begin{widetext}
\begin{align}
\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)
&= 
\frac{2}{\hbar\xi(\go)}
\left(\mu^{\mathrm{abxx}}(\omega)\cos^{2}\alpha + 
\mu^{\mathrm{abyy}}(\omega)\sin^{2}\alpha + 
\mu^{\mathrm{abxy}}(\omega)\sin 2\alpha\right)
\label{eq:vab-aw}
\end{align}
\end{widetext}
as for the structures chosen in this article,
$\xi^{\mathrm{xx}}(\go)=\xi^{\mathrm{yy}}(\go)\equiv\xi(\go)$, and $\xi^{\mathrm{xy}}(\go)=0$.
Now, we formalize our two options for $\calv^{\rma\rmb}(\go)$.

@@

Analyzing the SVI, Eq. \eqref{eq:vab-aw}, we define the magnitude of
the electron velocity in plane
with the spin polarized along the $\mathrm{b}$ direction as
\begin{equation}
\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)
\equiv
\sqrt{
[\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)]^{2}\ +
[\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)]^{2}\ 
}, 
\label{eq:vs-mag}
\end{equation}
and define the angle at which the velocity is directed on the $xy$ plane as
\begin{equation}
\gamma_{\gs^\mathrm{b}} (\omega,\alpha)
=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)} \right)
.
\label{eq:gamma-ang}
\end{equation}
% where this angle is measured in the counter-clockwise direction from the
% positive $x$. 
We also define two special angles
\begin{equation}
\gamma_{\gs^\mathrm{b}}^\parallel(\omega,\alpha) = \alpha, 
\label{eq:gamma-par} 
\end{equation}
and
\begin{equation}
\gamma_{\gs^\mathrm{b}}^\perp(\omega,\alpha) = \alpha \pm 90^{\circ},
\label{eq:gamma-perp}
\end{equation}
corresponding to the electron velocity being parallel or perpendicular
the incoming polarization, 
respectively. The subscript $\gs^\rmb$ denotes the spin along $\rmb$.

@@

\begin{table}[t]
\center
\begin{tabular}{ccccc}\\
\hline
\quad Layer \quad & \quad Atom \qquad & \multicolumn{3}{c}{Position (\AA)} \\
\cline{3-5}
\quad No.   \quad & \quad type \qquad & $x$ & $y$ & $z$  \\
\hline
1 & H & -0.61516 & -1.77416 &  0.73196 \\
1 & H &  0.61518 &  0.35514 &  0.73175 \\
2 & C & -0.61516 & -1.77264 & -0.49138 \\
2 & C & -0.61516 & -0.35600 & -0.72316 \\
2 & C &  0.61516 &  0.35763 & -0.49087 \\
\hline
\end{tabular}

\caption{Unit cell of Up structure. Layer division, atom types and
positions for the Up structure. The structure unit cell was divided in
two layers corresponding to hydrogen and carbon atoms.The corresponding layer
atom position is depicted in Fig. \ref{fig:up-struc} with the corresponding
number of layer.}
\label{tab:up-unitcell}
\end{table}
% 
% 
\begin{table}[t]
\center
\begin{tabular}{ccccc}\\
\hline
\quad Layer \quad & \quad Atom \qquad & \multicolumn{3}{c}{Position (\AA)} \\
\cline{3-5}
\quad No.   \quad & \quad type \qquad & $x$ & $y$ & $z$  \\
\hline
1 & H &  -0.61516 &  -1.42140 & \ 1.47237 \\
2 & C &  -0.61516 &  -1.73300 & \ 0.39631 \\
3 & C & \ 0.61516 & \ 1.73300 & \ 0.15807 \\
4 & C & \ 0.61516 & \ 0.42201 &  -0.15814 \\
5 & C &  -0.61516 &  -0.37396 &  -0.39632 \\
6 & H &  -0.61516 &  -0.68566 &  -1.47237 \\
\hline
\end{tabular}

\caption{Unit cell of Alt structure. Layer division, atom types and
positions for the Alt structure. The structure unit cell was divided in
six layers corresponding each one to atoms in different $z$ positions. The
corresponding layer atom position is depicted in Fig. \ref{fig:alt-struc} with
the corresponding number of layer.}
\label{tab:alt-unitcell}
\end{table}

We preset the results of $\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)$ and
$\mathcal{V}_{\mathrm{a}}(\omega,\alpha)$ for the C$_{16}$H$_{8}$-up  and
C$_{16}$H$_{8}$-alt structures being both noncentrosymmetric semi-infinite 2D
carbon systems with 50\% hydrogenation in different arrangements. We recall
that the Up structure has hydrogen atoms only on the upper side of the carbon
sheet while the Alt structure has alternating hydrogen atoms on the upper and
bottom sides. Also we take the hexagonal carbon lattice to be on the $xy$ plane
for both structures, and the carbon-hydrogen bonds on the perpendicular $xz$
plane, as depicted in Figs. \ref{fig:up-struc} and \ref{fig:alt-struc}. The
coordinates for the Up and Alt unit cells of the structures are presented in
Tables \ref{tab:up-unitcell} and \ref{tab:alt-unitcell}.

We calculated the self-consistent ground state and the Kohn-Sham states using
density functional theory in the local density approximation (DFT-LDA) with a
planewave basis using the ABINIT code \cite{gonzeCPC09}.
% 
We used Hartwigsen-Goedecker-Hutter (HGH) relativistic separable dual-space
Gaussian pseudopotentials \cite{hartwigsenPRB98} including the spin-orbit
interaction needed to calculate $\mu^{\mathrm{abcd}}(\omega,\alpha)$ presented
in Eq. \eqref{eq:mu}.
% 
The convergence parameters for the calculations of our results corresponding to
the Up and Alt structures are cutoff energies of 40\,Ha and 65\,Ha, resulting
in LDA energy band gaps of 0.084\,eV and 0.718\,eV, respectively. The energy
eigenvalues and matrix elements for the Up and Alt structures were calculated
using 12802 $\mathbf{k}$ points and 14452 $\mathbf{k}$ points in the IBZ to
integrate $\mu^{\mathrm{abcd}}(\go)$ and $\xi^{\rma\rmb}(\go)$ using the
linearized analytic tetrahedron method (LATM).\cite{nastosPRB07} We neglect the
anomalous velocity term $\hbar(\bfgs\times\nabla V)/4m^2c^2$, where $V$ is the
crystal potential, in $\hat\bfv$ of Eq.~\eqref{z.1}, as this term is known to
give small contribution to PSC.\cite{bhatPRL05} Therefore,
$[\hat\bfv,\hat\bfS]=0$ and Eq.~\eqref{z.1} reduces to $\hat K^{\rma\rmb}=\hat
v^\rma \hat S^\rmb=\hat S^\rmb \hat v^\rma$. Finally, the prime in the sum of
Eq.~\eqref{eq:mu} is restricted to quasidegenerated conduction bands $c$ and
$c'$ that are closer than 30 meV, where this value is both a typical laser-
pulse energy width and the room-temperature energy.\cite{nastosPRB07}
% 
% It is known that the DFT with LDA approximation misestimates the energy band
% gap of semiconductors. To correct this one has to include the many-body
% interaction using i.e. the so-called GW approximation but this technique has a
% very high computational cost and is out of the scope in this paper.
% Nevertheless, DFT still remains a mainstream tool for calculating diverse
% properties derived from the electronic band structure.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%% Res: mu & V^{ab} comparison %%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{plots/2-vab-str-comp}
\caption{$\mathcal{V}^{\mathrm{ab}}(\go,\ga)$ 
  vs. $\hbar\go$, for the values of $\ga$ that maximize the signal. The low
  energy regions for the Alt and Up structures start at the energy gap of each
  structure, while in the high energy regions, the values of
  $\mathcal{V}^{\mathrm{ab}}(\go,\ga)$ are also very large. The bottom panel
  corresponds to CdSe and GaAs. }
\label{fig:vab-str-comp}
\end{figure}
